---
title: "Research Unplugged: Introduction to R & RStudio"
author: |
  | Sam J. Vizcaino Vickers
  | Psychology Technician 
  | Department of Psychology, Wrexham University
  |
  | \faEnvelopeO \  \texttt{sam.vickers@wrexham.ac.uk}
  | \faEnvelopeO \  \texttt{psychology.technician@wrexham.ac.uk}
  | \faEnvelopeO \  \texttt{}
fontsize: 8pt
output:
 beamer_presentation:
  keep_tex: false
  fonttheme: "serif"
  includes:
   in_header: preamble2.tex
editor_options: 
  chunk_output_type: console
csl: apa.csl
---

```{r, include = FALSE}
# Package Setup --------------
library(MASS)
library(tidyverse)
```

# What is R?

- Open-source language designed for **statistical computing, analysis and open science.**

- Extensive Package Ecosystem: features over 20,000 packages (via [CRAN](https://cran.r-project.org/)) that allow users to perform several different statistical models and visualisations.

- Renowned for its ability to produce high quality graphics and interactive web apps

# What is RStudio?

- It is an integrated development environment (IDE)
- Built for Data Science
- Workspace for Scripting, academic Manuscripts, Presentations, and more...

# Alternatives?

- What about alternative programs such Python, SPSS, SAS, Stata, Minitab, Matlab, Maple, Mathematica, Tableau, Excel, SQL and many others) come down to 3 inter-related factors:

  - It is immensely powerful

  - Open source

  - Very and increasingly widely used

- Positron (IDE)

  - New upcoming IDE

  - Worth it?

# Essential Concepts

- Data Wrangling

- Data Visualisation

- Quarto

- Statistics

# What are we going to learn today

- *Guided tour of RStudio*

  - Calling R packages, functions, variables etc.

- *First Steps in R*

- Basic Data visualisations

# Histograms

```{r, echo=FALSE, warning=FALSE}
ggplot(anorexia, 
       mapping = aes(x = Prewt, fill = )) +
  geom_histogram(binwidth= 1.80)
```

# Frequency Polygons

```{r, echo=FALSE, warning=FALSE}
ggplot(anorexia, 
       mapping = aes(x = Prewt)
) + geom_freqpoly(binwidth = 2.54, colour = "red")
```

# Density Plots

```{r, echo=FALSE, warning=FALSE}
ggplot(anorexia, 
       mapping = aes(x = Prewt)
) + geom_density()
```

# Bar Plots

```{r, echo=FALSE, warning=FALSE}
ggplot(anorexia, 
       mapping = aes(x = Treat, fill = Treat)
       ) + geom_bar()
```

# Scatter Plot

```{r, echo=FALSE, warning=FALSE, message=FALSE}
ggplot(mtcars, aes(x = wt, y = mpg)) + 
  geom_point(alpha = 0.5) + 
  geom_smooth(method = "lm", se = FALSE, colour = "red") + 
  labs(
    title = "Car Weight vs Fuel Efficiency", 
    x = "Weight (1,000 lbs)", 
    y = "Miles Per Gallon (MPG)"
  ) + 
  theme_minimal()

```

# Growth Curve Plots

```{r, echo=FALSE, warning=FALSE}
ggplot(ChickWeight, aes(x = Time, y = weight, group = Chick, color = Diet)) +
  geom_line(alpha = 0.5) +
  geom_point() +
  theme_minimal()
```

# Facet Plots

```{r, echo=FALSE, warning=FALSE, message = FALSE}
ggplot(ChickWeight, aes(x = Time, y = weight, group = Chick, colour = Diet)) +
  geom_line() +
  geom_point() +
  facet_wrap(~ Chick) +
  labs(x = "Time", y = "weight") +
  theme_bw()
```

# Psychology Students

For any psychology Students who wish to learn more about R and RStudio then do book a session with us on the [Booking System](https://outlook.office.com/book/PsychTechWrexham@MAILGLYNDWRAC.onmicrosoft.com/?ismsaljsauthenabled){.uri}.
